Observing Economic Growth in Unrecognized States with Nighttime Light \thanks{I thank Elizaveta Kugaevskaia for her superb research assistance on this project. Financial support from Japan Society for the Promotion of Science (JSPS Kakenhi Grant Number 17K13730) is gratefully acknowledged.}

### 2019-03-19 paper/paper.tex
Revise the introduction #54
Improve the choice of words and phrases. Add the last paragraph on the structure of the paper. Add cross-references to each section

\section{Introduction}
% Introducing frozen conflicts
Back in 1991, the Soviet Union broke up into 15 of its constituent republics. 
Immediately after these republics became independent as sovereign states, wars of secession broke out in three of them: Azerbaijan, Georgia, and Moldova. 
Secessionists, consisting of minority ethnic groups in each republic, won the wars and declared the independence of their territory: Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova. 
However, not a single country around the world recognized them as sovereign states. 
These four ``states'' remain internationally unrecognized until today.%
\footnote{%
	Since 2008, Russia and a few other countries have recognized Abkhazia and South Ossetia as sovereign states, though.
}

This paper attempts to estimate the impact of being unrecognized as a sovereign state on economic activities in these four territories of the former Soviet Union. 
Although we focus only on four cases of unrecognized states, there are several other cases of unrecognized states.
Somaliland in northwest Somalia has not been recognized by any country since 1991. 
Taiwan and Northern Cyprus have been recognized only by a few countries since the 1970s. 
Since 2008, Kosovo has been recognized as independent by many countries, but not by the others including its parent country Serbia.
More recently, the 2014 crisis in Ukraine resulted in the creation of two unrecognized states in its eastern area: Luhansk and Donetsk.
It is therefore important to understand the consequences of being unrecognized as sovereign states because not an ignorable number of people now live in such pseudo states today.%
\footnote{
	There were more short-lived cases of unrecognized states: Katanga in Zaire (Democratic Republic of Congo) during 1960--63, Republika Srpska Krajina in Croatia during 1991--95, Republika Srpska in Bosnia-Herzegovina during 1992--95, and Chechnya in Russia during 1996-99.
	}

@@

This study contributes to four strands of the literature. 
First, the recent literature on the role of states in economic development focuses on the impacts of state capacity (Besley and Persson 2011, Acemoglu et al.\ 2015). 
However, state capacity is only one side of sovereign states. 
The other side is international recognition, which has been ignored by economists. 
This study fills this gap in the literature of economics.

Second, the issue of unrecognized states has been ignored by the emerging literature on the impacts of civil conflicts%
\footnote{
	See Blattman and Miguel (2010:Section 4) and Bauer et al.\ (2016) for surveys.
}  
even though the lack of international recognition is one of the likely consequences from those civil wars in which one side attempts to secede from the parent country. 
By filling this gap in the literature, our study makes two contributions.

The literature finds that output loss due to civil conflicts tends to be temporary (Davis and Weinstein 2002; Miguel and Roland 2011; Cerra and Saxena 2008).%
\footnote{
	Mueller (2012), commenting on the methodology in Cerra and Saxena (2008), shows that output persistently drops after the \textit{beginning} of civil conflicts by over 15\%.
} 
These findings, however, may suffer from sample selection bias, as pointed out by Blattman and Miguel (2010): the data is unavailable for those countries that severely suffer from civil conflicts.   
Unrecognized states are an example of the out-of-the-sample cases of post-conflict situations. 
Consequently, understanding the consequences of unrecognized states helps us paint the full picture of the impacts of civil conflicts.

@@

Third, our study relates to the political science literature on unrecognized states (King 2001, Lunch 2002, Kolsto 2006, O'Loughlin et al.\ 2014, Smolnik 2016, among others). 
In particular, a recent strand of this literature points out that seeking international recognition creates an incentive for politicians in unrecognized states to promote state-building and democratic institutions (Caspersen 2008, 2011; Blakkisrud and Kolsto 2011; Kolsto and Blakkisrud 2012).
Our study complements this literature by uncovering economic consequences of non-recognition which can be interpreted as an indirect impact via state-building and democracy.%
\footnote{
	Witmer and O'Loughlin (2011) use nighttime light data to gauge the impact of wars in towns in Georgia and South Ossetia. Our study differs in the coverage of study areas and in the recovery of economic growth rates from light data. 
}

@@

@@

%%%%%%%%%%%% THEORY %%%%%%%%%%%%
\section{Economic Consequences of Non-recognition}
The failure to be recognized as a sovereign state by international community may adversely affect economic development for at least two reasons.
First, non-recognition may discourage foreign direct investment (Kolsto 2006:729). 
There is uncertainty over whether legal contracts in unrecognized states are internationally binding. 
Investment in unrecognized states could also deny access to markets in their parent countries.
Second, the military expenditure may crowd out government investments in public goods.%
\footnote{
	In Nagorno-Karabakh, 65 out of 1,000 inhabitants are soldiers, surpassing almost all countries around the world (International Crisis Group 2005:9). 
	As of 2000, Transnistria had 5,000 to 10,000 soldiers; South Ossetia 2,000; Abkhazia 5,000 (International Institute for Strategic Studies 2000).
}
Parent countries can anytime invade unrecognized states to restore their control over the lost territory, as Georgia did to South Ossetia in 2008.

@@

\section{Data}
This section describes the datasets we use to recover economic growth rates in the four unrecognized states from satellite images of nighttime light.
\subsection{Real GDP}
We use the Penn World Table 9.0 (Freenstra et al.\ 2015) as the data source for real GDP. 
Expenditure-side real GDP at chained PPPs in 2011 US dollars (the variable \textit{rgdpe}) is used because the variable is meant for comparing ``living standards across countries and across years'' (Freenstra et al.\ 2015, Table 1).

@@

\section{Methodology}
% Model of real GDP
Following Henderson et al.\ (2012), we model real GDP in territory $i$ in year $t$, denoted by $y_{i,t}$, as follows:
\begin{equation}\label{gdp}
\ln y_{i,t} = \beta \ln L_{i,t} + \mu_i + \eta_t + \varepsilon_{i,t},
\end{equation}
where $L_{i,t}$ is nighttime light intensity in territory $i$ in year $t$, $\mu_i$ the territory fixed effect, $\eta_t$ the year fixed effect, and $\varepsilon_{i,t}$ the error term.

@@

\section{Results}
\subsection{Fixed Effects Estimation Results}
Table \ref{estimates} reports the results from estimating equation (\ref{gdp}). 
The elasticity of real GDP with respect to nighttime light intensity is 0.35 (significantly different from zero at the 1\% level).
This estimate is largely comparable to what the literature has found.%
\footnote{
	Henderson et al.\ (2012:Table 2) report the elasticity of 0.28 for the sample period of 1992--2008. Hodler and Raschky (2014:Appendix B) find the elasticity of 0.39 for subnational regions across the world for 1992--2009. Storeygard (2016:Table 1) obtains the elasticity of 0.25 for Chinese cities and prefectures for 1990--2005. 
}   
Year fixed effect estimates are also reported in Table \ref{estimates} as these values will be used to predict annual real GDP growth for unrecognized states. 
Note that the adjusted R-squared is 0.99. 
The model represented in equation (\ref{gdp}) explains almost all the variations in real GDP across countries and over time.

@@

\section{Conclusions}
% Discuss external validity
% Future research: precise border data and RD design

>>>>

\section{Introduction}
% Introducing frozen conflicts
Back in 1991, the Soviet Union broke up into fifteen of its constituent republics. 
Immediately after, wars of secession broke out in three of these newly independent states: Azerbaijan, Georgia, and Moldova.
Secessionists, consisting of minority ethnic groups in each republic, succeeded in taking over their territories and declared their independence from their parent states. 
These breakaway territories include Nagorno-Karabakh in Azerbaijan, Abkhazia and South Ossetia in Georgia, and Transnistria in Moldova (Figure \ref{map}). 
However, not a single country around the world recognized them as sovereign states. 
These four ``states'' remain internationally unrecognized until today.%
\footnote{%
	Since 2008, Russia and a few other countries have recognized Abkhazia and South Ossetia as sovereign states, though.
}

This paper attempts to estimate economic growth rates in these four unrecognized states of the former Soviet Union, with the aim of understanding the impact of non-recognition as a sovereign state on economic activities.

Unrecognized states are not so uncommon. 
Somaliland in northwest Somalia has not been recognized by any country since 1991. 
Taiwan and Northern Cyprus have been recognized only by a few countries since the 1970s. 
Since 2008, Kosovo has been recognized as independent by many countries, but not by the others including its parent country Serbia.
More recently, the 2014 crisis in Ukraine resulted in the creation of two unrecognized states in its eastern area: Luhansk and Donetsk.
It is therefore important to understand the consequences of being unrecognized as sovereign states because not an ignorable number of people now live in such pseudo states today.%
\footnote{
	There were also short-lived unrecognized states in the past: Katanga in Zaire (today's Democratic Republic of Congo) during 1960--63, Republika Srpska Krajina in Croatia during 1991--95, Republika Srpska in Bosnia-Herzegovina during 1992--95, and Chechnya in Russia during 1996--99.
	}

@@

This study relates to four strands of the literature. 
First, the recent literature on the role of states in economic development focuses on the impacts of state capacity (Besley and Persson 2011, Acemoglu et al.\ 2015). 
However, state capacity is only one side of sovereign states. 
The other side is international recognition, which has been ignored by economists. 
This study fills this gap in the literature.

Second, the issue of unrecognized states has received no attention in the emerging literature on the impacts of conflicts.%
\footnote{
	See Blattman and Miguel (2010:Section 4) and Bauer et al.\ (2016) for surveys.
}  
However, the lack of international recognition is one of the likely consequences from those civil wars in which one side attempts to secede from the parent country. 
By filling this gap in the literature, our study makes two contributions.

The literature finds that output loss due to conflicts tends to be temporary (Davis and Weinstein 2002; Miguel and Roland 2011; Cerra and Saxena 2008).%
\footnote{
	Mueller (2012), commenting on the methodology in Cerra and Saxena (2008), shows that output persistently drops after the \textit{beginning} of civil conflicts by over 15\%.
} 
These findings, however, may suffer from sample selection bias, as pointed out by Blattman and Miguel (2010): the data is unavailable for those countries that severely suffer from civil conflicts.   
Unrecognized states are an example of the out-of-the-sample cases of post-conflict situations. 
Consequently, understanding the consequences of unrecognized states will help us paint the full picture of the impacts of civil conflicts.

@@

The third literature that our study relates to is the political science literature on unrecognized states (King 2001, Lunch 2002, Kolsto 2006, O'Loughlin et al.\ 2014, Smolnik 2016, among others). 
In particular, a recent strand of this literature points out that seeking international recognition creates an incentive for politicians in unrecognized states to promote state-building and democratic institutions (Caspersen 2008, 2011; Blakkisrud and Kolsto 2011; Kolsto and Blakkisrud 2012).
Our study complements this literature by uncovering economic consequences of non-recognition which can be interpreted as an indirect impact via state-building and democracy.%
\footnote{
	Witmer and O'Loughlin (2011) also use nighttime light data to gauge the impact of wars in towns in Georgia and South Ossetia. Our study differs in the coverage of study areas and in the recovery of economic growth rates from light data. 
}

@@

The rest of the paper is organized as follows. Section \ref{background} overviews the issue of unrecognized states in the former Soviet Union. Section \ref{theory} provides theoretical arguments on how non-recognition as a sovereign state can affect economic activities. Sections \ref{data} and \ref{methodology} introduce the datasets and methodology that we employ to uncover economic growth rates in unrecognized states. Section \ref{results} reports our findings. The last section concludes with discussions on the applicability of our findings to other unrecognized states.

@@

%%%%%%%%%%%% THEORY %%%%%%%%%%%%
\section{Economic Consequences of Non-recognition}\label{theory}
The failure to be recognized as a sovereign state by international community may adversely affect economic development for at least two reasons.
First, non-recognition may discourage foreign direct investment (Kolsto 2006:729). 
There is uncertainty over whether legal contracts in unrecognized states are internationally binding. 
Investment in unrecognized states could also deny access to markets in their parent countries.
Second, the military expenditure may crowd out government investments in public goods.%
\footnote{
	In Nagorno-Karabakh, 65 out of 1,000 inhabitants are soldiers, surpassing almost all countries around the world (International Crisis Group 2005:9). 
	As of 2000, Transnistria had 5,000 to 10,000 soldiers; South Ossetia 2,000; Abkhazia 5,000 (International Institute for Strategic Studies 2000).
}
Parent countries can anytime invade unrecognized states to restore their control over the lost territory, as Georgia did to South Ossetia in 2008.

@@

\section{Data}\label{data}
This section describes the datasets we use to recover economic growth rates in the four unrecognized states from satellite images of nighttime light.
\subsection{Real GDP}
We use the Penn World Table 9.0 (Freenstra et al.\ 2015) as the data source for real GDP. 
Expenditure-side real GDP at chained PPPs in 2011 US dollars (the variable \textit{rgdpe}) is used because the variable is meant for comparing ``living standards across countries and across years'' (Freenstra et al.\ 2015, Table 1).

@@

\section{Methodology}\label{methodology}
% Model of real GDP
Following Henderson et al.\ (2012), we model real GDP in territory $i$ in year $t$, denoted by $y_{i,t}$, as follows:
\begin{equation}\label{gdp}
\ln y_{i,t} = \beta \ln L_{i,t} + \mu_i + \eta_t + \varepsilon_{i,t},
\end{equation}
where $L_{i,t}$ is nighttime light intensity in territory $i$ in year $t$, $\mu_i$ the territory fixed effect, $\eta_t$ the year fixed effect, and $\varepsilon_{i,t}$ the error term.

@@

\section{Results}\label{results}
\subsection{Fixed Effects Estimation Results}
Table \ref{estimates} reports the results from estimating equation (\ref{gdp}). 
The elasticity of real GDP with respect to nighttime light intensity is 0.35 (significantly different from zero at the 1\% level).
This estimate is largely comparable to what the literature has found.%
\footnote{
	Henderson et al.\ (2012:Table 2) report the elasticity of 0.28 for the sample period of 1992--2008. Hodler and Raschky (2014:Appendix B) find the elasticity of 0.39 for subnational regions across the world for 1992--2009. Storeygard (2016:Table 1) obtains the elasticity of 0.25 for Chinese cities and prefectures for 1990--2005. 
}   
Year fixed effect estimates are also reported in Table \ref{estimates} as these values will be used to predict annual real GDP growth for unrecognized states. 
Note that the adjusted R-squared is 0.99. 
The model represented in equation (\ref{gdp}) explains almost all the variations in real GDP across countries and over time.

@@

\section{Conclusions}\label{conclusions}
% Discuss external validity
% Future research: precise border data and RD design
